---
title: "Use EpiControl with compartmental models"
author: "Sandor Beregi"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  eval = TRUE,
  warning = FALSE,
  message = FALSE,
  collapse = TRUE,
  comment = "#>"
)
```

## Overview

We will use the MPC algorithm with an SIRDS model, where we apply
interventions reducing the transmission rate. Projections to evaluate
the cost function are still made using a renewal model.

### Model overview

## SIRDS compartmental model

We will implement a discrete-time, stochastic **SIRDS** compartmental
model with **two infectious compartments**:

-   **S**: Susceptible
-   **IR**: Infectious individuals who will **recover**
-   **ID**: Infectious individuals who will **die**
-   **R**: Recovered
-   **D**: Dead

The total infectious population is $$
I = I_R + I_D
$$ and the (living) population size used for transmission is $$
N = S + I + R.
$$

### Force of infection

Transmission is governed by the “force of infection” (FOI) which is
computed as `beta * S / N` and then applied via `event(I, FOI)` to
generate new infections.

### Stochastic (or event-based) transitions

Flows between compartments are generated via helper functions:

-   `event(x, rate)` draws the number transitioning out of a compartment
    size `x` at a given per-step rate (or hazard) over `dt`.
-   `branch(x, p)` splits an event count `x` into two outcomes with
    probability `p` (here used to allocate infections into fatal vs
    non-fatal infections).

At each time step:

1.  **Infections**:\
    $$
    S \rightarrow I
    $$ computed as `Infections <- event(I, FOI)`.

2.  **Outcome split (branching according to the infection fatality rate
    (IFR))**: each infection is assigned to

    -   **fatal infection** with probability $IFR$: $S \rightarrow I_D$
    -   **recovering infection** with probability $1-IFR$:
        $S \rightarrow I_R$

    implemented as `ID_and_IR <- branch(Infections, IFR)` where:

    -   `ID_and_IR[1]` = new $I_D$
    -   `ID_and_IR[2]` = new $I_R$

3.  **Recovery**: $$
    I_R \rightarrow R
    $$ at rate `recov_rate`.

4.  **Death**: $$
    I_D \rightarrow D
    $$ at rate `death_rate`.

5.  **Waning immunity**: $$
    R \rightarrow S
    $$ at rate `waning_rate`.

### Recorded event outputs

Based on the above, we will define an `"events"` attribute to the model
update so simulations can store incident flows and use them for control
purposes:

-   `Infections`
-   `New_IR`, `New_ID`
-   `Recoveries`
-   `Deaths`
-   `Losing_immunity`

This is also useful for plotting incidence time series (e.g., daily
infections or deaths) separately from compartment sizes.

### State update equations

Let $\Delta X$ denote the change in compartment $X$ over the time step.
The model updates are:

$$
\Delta S  = -\text{Infections} + \text{LosingImmunity}
$$ $$
\Delta I_R = \text{New_IR} - \text{Recoveries}
$$ $$
\Delta I_D = \text{New_ID} - \text{Deaths}
$$ $$
\Delta R  = \text{Recoveries} - \text{LosingImmunity}
$$ $$
\Delta D  = \text{Deaths}
$$

where each flow is computed using `event()` / `branch()` as described
above.

Initialisation:

```{r setup}
library(EpiControl)
library(VGAM)
library(parallel)
library(pbapply)
library(zoo)  # For rolling sum operations
library(dplyr)
library(ggplot2)
```

```{r fix random seed}
set.seed(1)
```

Let's define model and surveillance noise parameters.

```{r define-parameters}
Epi_pars <- data.frame(
  Pathogen = c("COVID-19"),
  R0 = c(2.2),
  sigma = c(1/6.5), #recovery rate
  gamma_d = c(1/6.5), #death rate
  zeta= c(1/80), # rate of losing immunity
  gen_time = c(6.5), # for the renewal model used in projections
  gen_time_var = c(2.1), # for the renewal model used in projections
  CFR = c(0.0132), # case fatality ratio
  mortality_mean = c(10.0), # mean time from infection to death (for renewal model projections)
  mortality_var = c(1.1) # mean time from infection to death (for renewal model projections)
)

Noise_pars <- data.frame(
  repd_mean = 10.5,  # Reporting delay mean
  del_disp = 5.0,    # Reporting delay variance
  ur_mean = 0.3,     # Under-reporting mean
  ur_beta_a = 50.0   # Beta distribution alpha for under-reporting
)
```

```{r define sim pars}
Action_space <- data.frame(
  NPI = c("No restrictions", "Lockdown"),
  R_coeff = c(1.0, 0.3),
  R_beta_a = c(0.0, 5.0),
  cost_of_NPI = c(0.0, 0.15)
)

C_target <- 5000
D_target <- 12
r_trans_len <- 7

sim_settings <- list(
  ndays = 120 * 7, #simulation length
  start_day = 1,
  N = 5e6, # population size
  I0 = 10, # initial infections
  C_target = C_target, #target cases
  C_target_pen = C_target*1.5, #overshoot penalty threshold
  R_target = 1.0,
  D_target = D_target, #one way to get peaks at 400 is to increase this to 15
  D_target_pen = 50, #max death
  alpha = 1.3/C_target, #~proportional gain (regulates error in cases) covid
  #alpha = 3.25/C_target #~proportional gain (regulates error in cases) ebola
  alpha_d = 0*1.3/D_target,
  ovp = 5.0, #overshoot penalty
  dovp = 0*10.0, #death overshoot penalty
  gamma = 0.95, #discounting factor
  n_ens = 10L, #MC assembly size for 4
  sim_ens = 10L, #assembly size for full simulation
  rf = 14L, #days 14
  R_est_wind = 5L, #rf-2 #window for R estimation
  susceptibles = 0,
  pred_days = 21L,
  r_trans_steep = 0.5,  # Growth rate
  r_trans_len = r_trans_len,  # Number of days for the transition
  t0 = r_trans_len  / 2, # Midpoint of the transition
  pathogen = 1,
  susceptibles = 0,
  delay = 1,
  ur = 1,
  r_dir = 2
)
```

Let's create the compartments and define flows.

```{r create flow}
# Create the compartments

model_pars <- list(
  beta = Epi_pars$R0 * Epi_pars$sigma,
  recov_rate = Epi_pars$sigma,
  death_rate = Epi_pars$gamma_d,
  waning_rate = Epi_pars$zeta,
  IFR = Epi_pars$CFR
)

flow_SIRDS <- function(t, state, pars, dt) {
  with(as.list(state), {
    I <- IR + ID
    N <- S + I + R

    FOI <- pars$beta * S / N

    # Events
    Infections <- event(I, FOI)
    ID_and_IR <- branch(Infections, pars$IFR)
    Recoveries <- event(IR, pars$recov_rate)
    Deaths <- event(ID, pars$death_rate)
    Losing_immunity <- event(R, pars$waning_rate)

    delta <- c(
      S  = -Infections + Losing_immunity,
      IR =  ID_and_IR[2] - Recoveries,
      ID =  ID_and_IR[1] - Deaths,
      R  =  Recoveries - Losing_immunity,
      D  =  Deaths
    )

    attr(delta, "events") <- c(
      Infections      = Infections,
      New_IR          = ID_and_IR[2],
      New_ID          = ID_and_IR[1],
      Recoveries      = Recoveries,
      Deaths          = Deaths,
      Losing_immunity = Losing_immunity
    )
    delta
  })
}

# ---- wire it up ----
S <- new_compartment("S", 1000000)
IR <- new_compartment("IR", 40)
ID <- new_compartment("ID", 0)
R <- new_compartment("R", 0)
D <- new_compartment("D", 0)

model <- build_model(S, IR, ID, R, D) |> set_flows(flow_SIRDS)

```

Let's define the events/calculated variables (e.g. reproduction number
estimates) we'd like to export and track (either for control purposes or
for plotting later, otherwise only compartments and flows are recorded
by default.) We also set up episim_data_ens: the dataframe where we will
store our results.

```{r event names}

event_names <- c("Infections", "New_IR", "New_ID", "Recoveries","Deaths", "Losing_immunity")

track_R <- c("Re","Rew","Rest","R0est","policy","R_coeff","I","C","Lambda","Lambda_C")

comp_names  <- names(state(model))  # e.g. S IR ID R D

model$events

column_names <- c("day","sim_id", comp_names, event_names, track_R)

episim_data <- data.frame(matrix(0, nrow = (sim_settings$ndays), ncol = length(column_names)))
colnames(episim_data) <- column_names

episim_data$policy <- rep(1, sim_settings$ndays)
episim_data$days <- 1:(sim_settings$ndays)

st0 <- state(model)
episim_data[1, comp_names] <- as.numeric(st0[comp_names])
episim_data[1, "Re"] <- Epi_pars[1, "R0"]
episim_data[1, "Rew"] <- Epi_pars[1, "R0"]
episim_data[1, "R0est"] <- Epi_pars[1, "R0"]
episim_data[1, "Rest"] <- Epi_pars[1, "R0"]
episim_data[1, "I"] <- 1

episim_data_ens <- replicate(sim_settings$sim_ens, episim_data, simplify = FALSE)

for (ii in 1:sim_settings$sim_ens) {
  episim_data_ens[[ii]]$sim_id <- rep(ii, sim_settings$ndays)
}

```

Let's define the simulation settings

```{r sim settings}

episettings <- list(
  sim_function = Epi_MPC_run_comp_model,
  model = model,
  incidence_flow = "Infections",
  transimssion_par = "beta",
  event_names = event_names,
  reward_function = reward_fun_wd,
  R_estimator = R_estim,
  noise_par = Noise_pars,
  epi_par = Epi_pars,
  model_par = model_pars,
  actions = Action_space,
  updatepars <- function(pars, ctx) pars,
  pars_to_change <- character(0),
  sim_settings = sim_settings,
  parallel = FALSE
)

```

In the list above updatepars and pars_to_change define how interventions
change transmission. De default setting for this is null, i.e. if
unchanged, interventions will not change anything. This is not what we
want, therfore we will define our custom function where the transmission
parameter is reduced according to R_coeff.

```{r intervention function}

episettings$updatepars <- function(pars, ctx) {
  R0      <- ctx$epi_par[ctx$pathogen, "R0"]
  sigma   <- ctx$epi_par[ctx$pathogen, "sigma"]
  gamma_d <- ctx$epi_par[ctx$pathogen, "gamma_d"]

  dyn_beta <- R0 * ctx$Rcoeff * sigma

  modifyList(pars, list(
    beta       = dyn_beta,
    beta0      = R0 * sigma,
    recov_rate = sigma,
    death_rate = gamma_d,
    P          = 0.95
  ))
}

```

Now we run the simulation.

```{r run-sim, warning=FALSE, message=FALSE}
# Ensure actions$cost_of_NPI is numeric
Action_space$cost_of_NPI <- as.numeric(as.character(Action_space$cost_of_NPI))

# Verify numeric parameters in sim_settings
numeric_params <- c("alpha", "alpha_d", "ovp", "dovp", "C_target", "C_target_pen", "D_target", "D_target_pen", "pred_days", "sim_ens", "ndays", "N")
sim_settings[numeric_params] <- lapply(sim_settings[numeric_params], as.numeric)

#Set this up for parallel run

# Create and export cluster
Ncores <-  2  # Adjust cores as appropriate
cl <- makeCluster(Ncores)

clusterExport(cl, ls())
clusterExport(cl, c("reward_fun_wd", "Epi_pred_wd", "Epi_MPC_run_wd", "episim_data_ens", "episettings", "Action_space", "Epi_pars", "Noise_pars", "R_estim"))

parallel::clusterEvalQ(cl, {
  library(pbapply)
  library(VGAM)
  library(EpiControl)
})

episettings$parallel <- TRUE
episettings$cl <- cl

# Run the epicontrol function
results <- epicontrol(episim_data_ens, episettings)

# Stop the cluster after simulation
parallel::stopCluster(cl)

```

Plot the results

```{r plotting}

# Combine Simulation Results
combined_data <- do.call(rbind, results)

combined_data <- combined_data %>%
  group_by(sim_id, policy) %>%
  arrange(days) %>%
  mutate(group = cumsum(c(1, diff(days) != 1))) %>%
  ungroup()

combined_data <- combined_data %>%
  arrange(sim_id, days, policy)

policy_labels <- c("1" = "No intervention", "2" = "Lockdown")


ggplot(combined_data %>% filter(sim_id == 1)) +
  geom_line(data = subset(combined_data, sim_id != 1), aes(x = days, y = C, color = as.factor(sim_id)), alpha = 0.1) +
  geom_line(aes(x = days, y = C, color = factor(policy, labels = policy_labels), group = 1), alpha = 1.0) +
  geom_line(aes(x = days, y = C, color = factor(policy, labels = policy_labels), group = 1), alpha = 1.0, linewidth=0.25) +
  geom_hline(yintercept = C_target, linetype = "dashed", color = "blue", linewidth=0.25) +
  labs(x = "Days", y = "Reported cases", color = "Policy") +
  scale_color_manual(values = c("No intervention" = "chartreuse3", "Lockdown" = "red")) +
  guides(color = guide_legend(title = "Policy"))

```
