---
title: "Use EpiControl with compartmental models"
author: "Sandor Beregi"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  eval = TRUE,
  warning = FALSE,
  message = FALSE,
  collapse = TRUE,
  comment = "#>"
)
```

## Overview

We will use the MPC algorithm with an SVIR model (SIR with vaccination),
where we apply interventions reducing the transmission rate. Projections
to evaluate the cost function are still made using a renewal model.

## SVIR model description

We consider an event-based, stochastic compartmental **SVIR** epidemic
model with four population groups:

-   **S**: susceptible individuals
-   **V**: vaccinated individuals
-   **II**: infectious individuals
-   **R**: recovered individuals

The total population size is fixed:

$$
N = S(t) + II(t) + R(t) + V(t).
$$

The model assumes that susceptible individuals may either become
vaccinated or infected. Infectious individuals recover at a constant
per-capita rate. The vaccine provides full immunity (vaccinated
individuals never get infected).

We introduce the model equations with the equilvalent (deterministic)
ordinary differential equation system:

$$
\frac{dS}{dt} = - \beta \frac{S}{N} II  - \nu S
$$

$$
\frac{dII}{dt} = \beta \frac{S}{N} II - \gamma II
$$

$$
\frac{dR}{dt} = \gamma II
$$

$$
\frac{dV}{dt} = \nu S.
$$

This means that:

-   susceptible individuals leave $S$ through either infection or
    vaccination,
-   newly infected individuals enter $II$,
-   infectious individuals recover into $R$,
-   vaccinated individuals move into $V$.

and $\beta, \gamma$ and $\nu$ are the infection, recovery and
vaccination rates.

### Stochastic event interpretation

In our implementation, infection, vaccination, and recovery are
represented as event-driven transitions over each time step $dt$. For
susceptible individuals, infection and vaccination compete as
alternative events, while recovery acts on the infectious compartment.
The realised changes in each compartment are then aggregated into the
state update.

The model records the following event counts at each step:

-   **Infections**
-   **Vaccinations**
-   **Recoveries**

and the change of the compartment populations are given by

$$
\Delta S = -\text{Infections} - \text{Vaccinations}
$$ $$
\Delta V =  \text{Vaccinations}
$$ $$
\Delta II = \text{Infections} - \text{Recoveries}
$$

$$
\Delta R = \text{Recoveries}
$$

## Initialisation

```{r setup}
library(EpiControl)
library(VGAM)
library(parallel)
library(pbapply)
library(zoo)  # For rolling sum operations
library(dplyr)
library(ggplot2)
```

```{r fix random seed}
set.seed(1)
```

Let's define model and surveillance noise parameter accounting for
epresent reporting delays and incomplete ascertainment in observed case
counts. We model reported cases $C_t$ as a noisy observation of the
underlying infectious process $I_t$. To account for reporting delays,
the expected number of reported cases at time $t$ is constructed by
convolving past infectious counts with a delay distribution
`Gamma(1:ndays, shape = gen_time / gen_time_var, rate = 1 / gen_time_var)`.
The delayed count is then observed through a Poisson sampling step
`pois_input_c <- sum(episimdata[ii:1, 'I'] * Ydel[1:ii])` and
`C_del <- rpois(1, pois_input_c)` . Optional under-reporting is
incorporated using a beta-binomial observation model
`C <- VGAM::rbetabinom.ab(1, C_del, ur_beta_a, ur_beta_b)` with
`ur_beta_b <- (1 - ur_mean) / ur_mean * ur_beta_a`, applied either to
the delayed count or, when no delay is included, directly to $I_t$.

```{r define-parameters}
Epi_pars <- data.frame(
  Pathogen = c("COVID-19"),
  R0 = c(2.8),
  sigma = c(1/6.5), #recovery rate
  vaccination_rate = c(1/1000), #vaccination rate
  gen_time = c(6.5), # for the renewal model used in projections
  gen_time_var = c(6.5), # for the renewal model used in projections
  CFR = c(0.0132), # case fatality ratio
  mortality_mean = c(10.0), # mean time from infection to death (for renewal model projections)
  mortality_var = c(1.1) # mean time from infection to death (for renewal model projections)
)

Noise_pars <- data.frame(
  repd_mean = 10.5,  # Reporting delay mean
  del_disp = 5.0,    # Reporting delay variance
  ur_mean = 0.3,     # Under-reporting mean
  ur_beta_a = 50.0   # Beta distribution alpha for under-reporting
)
```

### Simulation/model projection parameters and cost function

We choose the optimal NPI every `rf = 14` days to optimise the following
cost function (defined by reward_fun_wd.R):
`alpha * abs(C - C_target) + ovp + NPI_cost` where `ovp` is the penalty
for cases going above `C_target_pen` We only define a target for daily
cases, the cost function requires a death target as well, but we will
set the corresponding coefficients to zero.

We project forward using a renewal branching process model for a time
window of `pred_days = 21 days`. We use an ensemble of `n_ens = 10`
simulations to derive the expected cost of each NPI in Action_space.

```{r define sim pars}
Action_space <- data.frame(
  NPI = c("No restrictions", "Social distancing", "Lockdown"),
  R_coeff = c(1.0, 0.5, 0.2),
  R_beta_a = c(0.0, 5.0, 5.0),
  cost_of_NPI = c(0.0, 0.03, 0.15)
)

C_target <- 1000
D_target <- 12
r_trans_len <- 7

sim_settings <- list(
  ndays = 120 * 7, #simulation length
  start_day = 1,
  N = 5e6, # population size
  I0 = 10, # initial infections
  C_target = C_target, #target cases
  C_target_pen = C_target*5, #overshoot penalty threshold
  R_target = 1.0,
  D_target = D_target, #one way to get peaks at 400 is to increase this to 15
  D_target_pen = 50, #max death
  alpha = 1.3/C_target, #~proportional gain (regulates error in cases) covid
  #alpha = 3.25/C_target #~proportional gain (regulates error in cases) ebola
  alpha_d = 0*1.3/D_target,
  ovp = 5.0, #overshoot penalty
  dovp = 0*10.0, #death overshoot penalty
  gamma = 0.95, #discounting factor
  n_ens = 10L, #MC assembly size for 4
  sim_ens = 10L, #assembly size for full simulation
  rf = 14L, #policy review frequency: every days 14
  R_est_wind = 5L, #rf-2 #window for R estimation
  susceptibles = 0,
  pred_days = 21L,
  r_trans_steep = 0.5,  # Growth rate
  r_trans_len = r_trans_len,  # Number of days for the transition
  t0 = r_trans_len  / 2, # Midpoint of the transition
  pathogen = 1,
  delay = 1,
  ur = 1,
  r_dir = 2
)
```

### Model definition

Let's create the compartments and define flows. For simple transitions
between compartments we use the 'event' function, whereas for competing
events (infection and vacciantion for susceptible individuals) we will
use the 'branch' function.

```{r create flow}
# Create the compartments

model_pars <- list(
  beta = Epi_pars$R0 * Epi_pars$sigma,
  recov_rate = Epi_pars$sigma,
  vac_rate = Epi_pars$vaccination_rate
)

flow_SVIR <- function(t, state, pars, dt) {
  with(as.list(state), {
    N <- S + II + R + V

    inf_rate <- pars$beta * II / N
    vac_rate <- pars$vac_rate

    # Events
    Inf_and_vac <- event(S, inf_rate+vac_rate)
    II_and_VV <- branch(Inf_and_vac, inf_rate/(inf_rate+vac_rate))
    Infections <- II_and_VV[1]
    Vaccinations <- II_and_VV[2]
    Recoveries <- event(II, pars$recov_rate)

    delta <- c(
      S  = -Infections - Vaccinations,
      II =  Infections - Recoveries,
      R  =  Recoveries,
      V  =  Vaccinations
    )

    attr(delta, "events") <- c(
      Infections      = Infections,
      Vaccinations    = Vaccinations,
      Recoveries      = Recoveries
    )
    delta
  })
}

# ---- wire it up ----
S <- new_compartment("S", 6000000)
V <- new_compartment("V", 0)
II <- new_compartment("II", 40)
R <- new_compartment("R", 0)

model <- build_model(S, V, II, R) |> set_flows(flow_SVIR)

```

Let's add the events/calculated variables (e.g. reproduction number
estimates) we'd like to export and track to restuls (either for control
purposes or for plotting later, otherwise only compartments and flows
are recorded by default.) We also set up episim_data_ens: the dataframe
where we will store our results.

```{r event names}

event_names <- c("Infections", "Vaccinations", "Recoveries")

track_R <- c("Re","Rew","Rest","R0est","policy","R_coeff","I","C","Lambda","Lambda_C")

comp_names  <- names(state(model))

model$events

column_names <- c("day","sim_id", comp_names, event_names, track_R)

episim_data <- data.frame(matrix(0, nrow = (sim_settings$ndays), ncol = length(column_names)))
colnames(episim_data) <- column_names

episim_data$policy <- rep(1, sim_settings$ndays)
episim_data$days <- 1:(sim_settings$ndays)

st0 <- state(model)
episim_data[1, comp_names] <- as.numeric(st0[comp_names])
episim_data[1, "Re"] <- Epi_pars[1, "R0"]
episim_data[1, "Rew"] <- Epi_pars[1, "R0"]
episim_data[1, "R0est"] <- Epi_pars[1, "R0"]
episim_data[1, "Rest"] <- Epi_pars[1, "R0"]
episim_data[1, "I"] <- 1

episim_data_ens <- replicate(sim_settings$sim_ens, episim_data, simplify = FALSE)

for (ii in 1:sim_settings$sim_ens) {
  episim_data_ens[[ii]]$sim_id <- rep(ii, sim_settings$ndays)
}

```

### Simulation

Let's define the simulation settings

```{r sim settings}

episettings <- list(
  sim_function = Epi_MPC_run_comp_model,
  model = model,
  incidence_flow = "Infections",
  transimssion_par = "beta",
  event_names = event_names,
  reward_function = reward_fun_wd,
  R_estimator = R_estim,
  noise_par = Noise_pars,
  epi_par = Epi_pars,
  model_par = model_pars,
  actions = Action_space,
  updatepars <- function(pars, ctx) pars,
  pars_to_change <- character(0),
  sim_settings = sim_settings,
  parallel = FALSE
)

```

In the list above updatepars and pars_to_change define how interventions
change transmission. De default setting for this is null, i.e. if
unchanged, interventions will not change anything. This is not what we
want, therefore we will define our custom function where the
transmission parameter is reduced according to R_coeff.

```{r intervention function}

episettings$updatepars <- function(pars, ctx) {
  R0      <- ctx$epi_par[ctx$pathogen, "R0"]
  sigma   <- ctx$epi_par[ctx$pathogen, "sigma"]

  dyn_beta <- R0 * ctx$Rcoeff * sigma

  modifyList(pars, list(
    beta       = dyn_beta
  ))
}

```

Now we run the simulation.

```{r run-sim, warning=FALSE, message=FALSE}
# Ensure actions$cost_of_NPI is numeric
Action_space$cost_of_NPI <- as.numeric(as.character(Action_space$cost_of_NPI))

# Verify numeric parameters in sim_settings
numeric_params <- c("alpha", "alpha_d", "ovp", "dovp", "C_target", "C_target_pen", "D_target", "D_target_pen", "pred_days", "sim_ens", "ndays", "N")
sim_settings[numeric_params] <- lapply(sim_settings[numeric_params], as.numeric)

#Set this up for parallel run

# Create and export cluster
Ncores <-  2  # Adjust cores as appropriate
cl <- makeCluster(Ncores)

clusterExport(cl, ls())
clusterExport(cl, c("reward_fun_wd", "Epi_pred_wd", "Epi_MPC_run_wd", "episim_data_ens", "episettings", "Action_space", "Epi_pars", "Noise_pars", "R_estim"))

parallel::clusterEvalQ(cl, {
  library(pbapply)
  library(VGAM)
  library(EpiControl)
})

episettings$parallel <- TRUE
episettings$cl <- cl

# Run the epicontrol function
results <- epicontrol(episim_data_ens, episettings)

# Stop the cluster after simulation
parallel::stopCluster(cl)

```

### Results

When we plot the results we observe that the controller aims to keep
daily incidence near the target level. We focus on the highlighted realisation, while also showing other epidemic realisations as grey curves to illustrate that, although stochasticity affects the exact trajectory, the overall behaviour remains similar. As immunity in the population
builds up, this can be achieved by periods of "social distancing" and no
intervention whilst the most stringent measure, "lockdown" is used only
once at the beginning when population immunity is low (below 20%). Once
herd immunity is reached no further intervention is applied. At this
point around 20% of the population has infection acquired immunity
(recovered) and another 45% is protected by the vaccine. The oscillations in incidence arise because the controller can only choose between three discrete intervention options and policies are reviewed every 14 days. As a result, it cannot continuously fine-tune transmission to maintain exactly 
R=1. In addition, reporting delays mean that control actions are based on lagged information, which can lead to responses being applied slightly late.

```{r plotting, fig.width=8, fig.height=7}

# Combine Simulation Results
combined_data <- do.call(rbind, results)

combined_data <- combined_data %>%
  group_by(sim_id, policy) %>%
  arrange(days) %>%
  mutate(group = cumsum(c(1, diff(days) != 1))) %>%
  ungroup()

combined_data <- combined_data %>%
  arrange(sim_id, days, policy)

# show the herd immunity threshold
R0 <- Epi_pars$R0
herd_immunity_threshold <- (1 - 1 / R0) * 100

sim1_data <- combined_data %>%
  filter(sim_id == 1) %>%
  mutate(
    immune_pct = (R + V) * 100 / (S + II + V + R)
  )

# first day herd immunity is reached in sim_id == 1
reach_day <- sim1_data %>%
  filter(immune_pct >= herd_immunity_threshold) %>%
  summarise(reach_day = min(days)) %>%
  pull(reach_day)

policy_labels <- c("1" = "No intervention", "2" = "Social distancing", "3" = "Lockdown")

library(patchwork)

top<-ggplot(combined_data %>% filter(sim_id == 1)) +
  geom_line(data = subset(combined_data, sim_id != 1), aes(x = days, y = C, color = as.factor(sim_id)), alpha = 0.1) +
  geom_line(aes(x = days, y = C, color = factor(policy, labels = policy_labels), group = 1), alpha = 1.0) +
  geom_line(aes(x = days, y = C, color = factor(policy, labels = policy_labels), group = 1), alpha = 1.0, linewidth=0.25) +
  geom_hline(yintercept = C_target, linetype = "dashed", color = "blue", linewidth=0.25) +
  annotate(
  "text",
  x = sim_settings$ndays * 0.9,
  y = C_target,
  label = "Target",
  vjust = -0.5,
  hjust = 0,
  size = 3
) +
  geom_vline(
    xintercept = reach_day,
    linetype = "dashed",
    colour = "black",
    linewidth = 0.5
  ) +
  labs(x = "Days", y = "Reported cases", color = "Policy") +
  scale_color_manual(values = c("No intervention" = "chartreuse3", "Social distancing" = "violet", "Lockdown" = "red")) +
  guides(color = guide_legend(title = "Policy"))

# Top panel
bottom <- ggplot(combined_data %>% filter(sim_id == 1), aes(x = days)) +
  geom_area(aes(y = (R+V)*100 / (S + II + V + R), fill = "V"), alpha = 0.8) +
  geom_area(aes(y = (R)*100 / (S + II + V + R), fill = "R"), alpha = 0.8) +
  labs(
    x = "Days",
    y = "Immune population %",
    fill = NULL
  ) +
  geom_hline(
    yintercept = herd_immunity_threshold,
    linetype = "dashed",
    colour = "black",
    linewidth = 0.4
  ) +
  geom_vline(
    xintercept = reach_day,
    linetype = "dashed",
    colour = "black",
    linewidth = 0.5
  ) +
annotate(
  "text",
  x = sim_settings$ndays * 0.6,
  y = herd_immunity_threshold,
  label = "Herd immunity threshold",
  vjust = -0.5,
  hjust = 0,
  size = 3
) +
  scale_fill_manual(
    values = c(
      "V" = "steelblue",
      "R" = "tomato"
    )
  ) +
  theme_minimal()

top/bottom

```

### Simulate with a higher vaccination rate

Let's see what happens if we vaccinate people 2x faster.

```{r change-vaccination-rate}
model_pars$vac_rate <- Epi_pars$vaccination_rate * 2

episim_data <- data.frame(matrix(0, nrow = (sim_settings$ndays), ncol = length(column_names)))
colnames(episim_data) <- column_names

episim_data$policy <- rep(1, sim_settings$ndays)
episim_data$days <- 1:(sim_settings$ndays)

st0 <- state(model)
episim_data[1, comp_names] <- as.numeric(st0[comp_names])
episim_data[1, "Re"] <- Epi_pars[1, "R0"]
episim_data[1, "Rew"] <- Epi_pars[1, "R0"]
episim_data[1, "R0est"] <- Epi_pars[1, "R0"]
episim_data[1, "Rest"] <- Epi_pars[1, "R0"]
episim_data[1, "I"] <- 1

episim_data_ens <- replicate(sim_settings$sim_ens, episim_data, simplify = FALSE)

for (ii in 1:sim_settings$sim_ens) {
  episim_data_ens[[ii]]$sim_id <- rep(ii, sim_settings$ndays)
}

cl <- makeCluster(Ncores)

clusterExport(cl, ls())
clusterExport(cl, c("reward_fun_wd", "Epi_pred_wd", "Epi_MPC_run_wd", "episim_data_ens", "episettings", "Action_space", "Epi_pars", "Noise_pars", "R_estim"))

parallel::clusterEvalQ(cl, {
  library(pbapply)
  library(VGAM)
  library(EpiControl)
})

episettings$cl <- cl
episettings$model_par <- model_pars

# Run the epicontrol function
results <- epicontrol(episim_data_ens, episettings)

# Stop the cluster after simulation
parallel::stopCluster(cl)


```

### Results

We can see that the accelerated vaccination programme leads to a shorter
outbreak and roughly a 50% reduction in the total number of infections.
With the higher vaccination rate, herd immunity is reached after 420
days rather than around 700, and by that point only around 12% of the
population has been infected.

Note that these simulations do not account for vaccination costs, the
cumulative burden of infections in the objective function, or any waning
of acquired immunity. Nevertheless, they demonstrate how EpiControl can
be used for this kind of scenario analysis.

```{r plotting2, fig.width=8, fig.height=7}

# Combine Simulation Results
combined_data <- do.call(rbind, results)

combined_data <- combined_data %>%
  group_by(sim_id, policy) %>%
  arrange(days) %>%
  mutate(group = cumsum(c(1, diff(days) != 1))) %>%
  ungroup()

combined_data <- combined_data %>%
  arrange(sim_id, days, policy)

sim1_data <- combined_data %>%
  filter(sim_id == 1) %>%
  mutate(
    immune_pct = (R + V) * 100 / (S + II + V + R)
  )

# first day herd immunity is reached in sim_id == 1
reach_day <- sim1_data %>%
  filter(immune_pct >= herd_immunity_threshold) %>%
  summarise(reach_day = min(days)) %>%
  pull(reach_day)


policy_labels <- c("1" = "No intervention", "2" = "Social distancing", "3" = "Lockdown")

library(patchwork)

top<-ggplot(combined_data %>% filter(sim_id == 1)) +
  geom_line(data = subset(combined_data, sim_id != 1), aes(x = days, y = C, color = as.factor(sim_id)), alpha = 0.1) +
  geom_line(aes(x = days, y = C, color = factor(policy, labels = policy_labels), group = 1), alpha = 1.0) +
  geom_line(aes(x = days, y = C, color = factor(policy, labels = policy_labels), group = 1), alpha = 1.0, linewidth=0.25) +
  geom_hline(yintercept = C_target, linetype = "dashed", color = "blue", linewidth=0.25) +
  annotate(
  "text",
  x = sim_settings$ndays * 0.9,
  y = C_target,
  label = "Target",
  vjust = -0.5,
  hjust = 0,
  size = 3
) +
  geom_vline(
    xintercept = reach_day,
    linetype = "dashed",
    colour = "black",
    linewidth = 0.5
  ) +
  labs(x = "Days", y = "Reported cases", color = "Policy") +
  scale_color_manual(values = c("No intervention" = "chartreuse3", "Social distancing" = "violet", "Lockdown" = "red")) +
  guides(color = guide_legend(title = "Policy"))

# Top panel
bottom <- ggplot(combined_data %>% filter(sim_id == 1), aes(x = days)) +
  geom_area(aes(y = (R+V)*100 / (S + II + V + R), fill = "V"), alpha = 0.8) +
  geom_area(aes(y = (R)*100 / (S + II + V + R), fill = "R"), alpha = 0.8) +
  labs(
    x = "Days",
    y = "Immune population %",
    fill = NULL
  ) +
  geom_hline(
    yintercept = herd_immunity_threshold,
    linetype = "dashed",
    colour = "black",
    linewidth = 0.4
  ) +
  geom_vline(
    xintercept = reach_day,
    linetype = "dashed",
    colour = "black",
    linewidth = 0.5
  ) +
annotate(
  "text",
  x = sim_settings$ndays * 0.6,
  y = herd_immunity_threshold,
  label = "Herd immunity threshold",
  vjust = -0.5,
  hjust = 0,
  size = 3
) +
  scale_fill_manual(
    values = c(
      "V" = "steelblue",
      "R" = "tomato"
    )
  ) +
  theme_minimal()

top/bottom

```
