---
title: "Use EpiControl with compartmental models"
author: "Sandor Beregi"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  eval = TRUE,
  warning = FALSE,
  message = FALSE,
  collapse = TRUE,
  comment = "#>"
)
```

## Overview

We will use the MPC algorithm with an SVIR model (SIR with vaccination),
where we apply interventions reducing the transmission rate. Projections
to evaluate the cost function are still made using a renewal model.

## SVIR model description

We consider an event-based, stochastic compartmental **SVIR** epidemic
model with four population groups:

-   **S**: susceptible individuals
-   **V**: vaccinated individuals
-   **II**: infectious individuals
-   **R**: recovered individuals

The total population size is fixed:

$$
N = S(t) + II(t) + R(t) + V(t).
$$

The model assumes that susceptible individuals may either become
vaccinated or infected. Infectious individuals recover at a constant
per-capita rate. The vaccine provides full immunity (vaccinated
individuals never get infected).

We introduce the model equations with the equilvalent (deterministic)
ordinary differential equation system:

$$
\frac{dS}{dt} = - \beta \frac{S}{N} II  - \nu S
$$

$$
\frac{dII}{dt} = \beta \frac{S}{N} II - \gamma II
$$

$$
\frac{dR}{dt} = \gamma II
$$

$$
\frac{dV}{dt} = \nu S.
$$

This means that:

-   susceptible individuals leave $S$ through either infection or
    vaccination,
-   newly infected individuals enter $II$,
-   infectious individuals recover into $R$,
-   vaccinated individuals move into $V$.

and $\beta, \gamma$ and $\nu$ are the infection, recovery and vaccination rates.

### Stochastic event interpretation

In our implementation, infection, vaccination, and recovery are
represented as event-driven transitions over each time step $dt$. For
susceptible individuals, infection and vaccination compete as
alternative events, while recovery acts on the infectious compartment.
The realised changes in each compartment are then aggregated into the
state update.

The model records the following event counts at each step:

-   **Infections**
-   **Vaccinations**
-   **Recoveries**

and the change of the compartment populations are given by

$$
\Delta S = -\text{Infections} - \text{Vaccinations}
$$ $$
\Delta V =  \text{Vaccinations}
$$ $$
\Delta II = \text{Infections} - \text{Recoveries}
$$

$$
\Delta R = \text{Recoveries}
$$

## Initialisation

```{r setup}
library(EpiControl)
library(VGAM)
library(parallel)
library(pbapply)
library(zoo)  # For rolling sum operations
library(dplyr)
library(ggplot2)
```

```{r fix random seed}
set.seed(1)
```

Let's define model and surveillance noise parameters.

```{r define-parameters}
Epi_pars <- data.frame(
  Pathogen = c("COVID-19"),
  R0 = c(2.8),
  sigma = c(1/6.5), #recovery rate
  vaccination_rate = c(1/1000), #vaccination rate
  gen_time = c(6.5), # for the renewal model used in projections
  gen_time_var = c(6.5), # for the renewal model used in projections
  CFR = c(0.0132), # case fatality ratio
  mortality_mean = c(10.0), # mean time from infection to death (for renewal model projections)
  mortality_var = c(1.1) # mean time from infection to death (for renewal model projections)
)

Noise_pars <- data.frame(
  repd_mean = 10.5,  # Reporting delay mean
  del_disp = 5.0,    # Reporting delay variance
  ur_mean = 0.3,     # Under-reporting mean
  ur_beta_a = 50.0   # Beta distribution alpha for under-reporting
)
```

```{r define sim pars}
Action_space <- data.frame(
  NPI = c("No restrictions", "Social distancing", "Lockdown"),
  R_coeff = c(1.0, 0.5, 0.2),
  R_beta_a = c(0.0, 5.0, 5.0),
  cost_of_NPI = c(0.0, 0.03, 0.15)
)

C_target <- 1000
D_target <- 12
r_trans_len <- 7

sim_settings <- list(
  ndays = 120 * 7, #simulation length
  start_day = 1,
  N = 5e6, # population size
  I0 = 10, # initial infections
  C_target = C_target, #target cases
  C_target_pen = C_target*5, #overshoot penalty threshold
  R_target = 1.0,
  D_target = D_target, #one way to get peaks at 400 is to increase this to 15
  D_target_pen = 50, #max death
  alpha = 1.3/C_target, #~proportional gain (regulates error in cases) covid
  #alpha = 3.25/C_target #~proportional gain (regulates error in cases) ebola
  alpha_d = 0*1.3/D_target,
  ovp = 5.0, #overshoot penalty
  dovp = 0*10.0, #death overshoot penalty
  gamma = 0.95, #discounting factor
  n_ens = 10L, #MC assembly size for 4
  sim_ens = 10L, #assembly size for full simulation
  rf = 14L, #days 14
  R_est_wind = 5L, #rf-2 #window for R estimation
  susceptibles = 0,
  pred_days = 21L,
  r_trans_steep = 0.5,  # Growth rate
  r_trans_len = r_trans_len,  # Number of days for the transition
  t0 = r_trans_len  / 2, # Midpoint of the transition
  pathogen = 1,
  susceptibles = 0,
  delay = 1,
  ur = 1,
  r_dir = 2
)
```

### Model definition

Let's create the compartments and define flows. For simple transitions
between compartments we use the 'event' function, whereas for competing
events (infection and vacciantion for susceptible individuals) we will
use the 'branch' function.

```{r create flow}
# Create the compartments

model_pars <- list(
  beta = Epi_pars$R0 * Epi_pars$sigma,
  recov_rate = Epi_pars$sigma,
  vac_rate = Epi_pars$vaccination_rate
)

flow_SVIR <- function(t, state, pars, dt) {
  with(as.list(state), {
    N <- S + II + R + V

    inf_rate <- pars$beta * II / N
    vac_rate <- pars$vac_rate

    # Events
    Inf_and_vac <- event(S, inf_rate+vac_rate)
    II_and_VV <- branch(Inf_and_vac, inf_rate/(inf_rate+vac_rate))
    Infections <- II_and_VV[1]
    Vaccinations <- II_and_VV[2]
    Recoveries <- event(II, pars$recov_rate)

    delta <- c(
      S  = -Infections - Vaccinations,
      II =  Infections - Recoveries,
      R  =  Recoveries,
      V  =  Vaccinations
    )

    attr(delta, "events") <- c(
      Infections      = Infections,
      Vaccinations    = Vaccinations,
      Recoveries      = Recoveries
    )
    delta
  })
}

# ---- wire it up ----
S <- new_compartment("S", 6000000)
V <- new_compartment("V", 0)
II <- new_compartment("II", 40)
R <- new_compartment("R", 0)

model <- build_model(S, V, II, R) |> set_flows(flow_SVIR)

```

Let's add the events/calculated variables (e.g. reproduction number
estimates) we'd like to export and track to restuls (either for control
purposes or for plotting later, otherwise only compartments and flows
are recorded by default.) We also set up episim_data_ens: the dataframe
where we will store our results.

```{r event names}

event_names <- c("Infections", "Vaccinations", "Recoveries")

track_R <- c("Re","Rew","Rest","R0est","policy","R_coeff","I","C","Lambda","Lambda_C")

comp_names  <- names(state(model))  # e.g. S IR ID R D

model$events

column_names <- c("day","sim_id", comp_names, event_names, track_R)

episim_data <- data.frame(matrix(0, nrow = (sim_settings$ndays), ncol = length(column_names)))
colnames(episim_data) <- column_names

episim_data$policy <- rep(1, sim_settings$ndays)
episim_data$days <- 1:(sim_settings$ndays)

st0 <- state(model)
episim_data[1, comp_names] <- as.numeric(st0[comp_names])
episim_data[1, "Re"] <- Epi_pars[1, "R0"]
episim_data[1, "Rew"] <- Epi_pars[1, "R0"]
episim_data[1, "R0est"] <- Epi_pars[1, "R0"]
episim_data[1, "Rest"] <- Epi_pars[1, "R0"]
episim_data[1, "I"] <- 1

episim_data_ens <- replicate(sim_settings$sim_ens, episim_data, simplify = FALSE)

for (ii in 1:sim_settings$sim_ens) {
  episim_data_ens[[ii]]$sim_id <- rep(ii, sim_settings$ndays)
}

```

### Simulation

Let's define the simulation settings

```{r sim settings}

episettings <- list(
  sim_function = Epi_MPC_run_comp_model,
  model = model,
  incidence_flow = "Infections",
  transimssion_par = "beta",
  event_names = event_names,
  reward_function = reward_fun_wd,
  R_estimator = R_estim,
  noise_par = Noise_pars,
  epi_par = Epi_pars,
  model_par = model_pars,
  actions = Action_space,
  updatepars <- function(pars, ctx) pars,
  pars_to_change <- character(0),
  sim_settings = sim_settings,
  parallel = FALSE
)

```

In the list above updatepars and pars_to_change define how interventions
change transmission. De default setting for this is null, i.e. if
unchanged, interventions will not change anything. This is not what we
want, therefore we will define our custom function where the
transmission parameter is reduced according to R_coeff.

```{r intervention function}

episettings$updatepars <- function(pars, ctx) {
  R0      <- ctx$epi_par[ctx$pathogen, "R0"]
  sigma   <- ctx$epi_par[ctx$pathogen, "sigma"]

  dyn_beta <- R0 * ctx$Rcoeff * sigma

  modifyList(pars, list(
    beta       = dyn_beta
  ))
}

```

Now we run the simulation.

```{r run-sim, warning=FALSE, message=FALSE}
# Ensure actions$cost_of_NPI is numeric
Action_space$cost_of_NPI <- as.numeric(as.character(Action_space$cost_of_NPI))

# Verify numeric parameters in sim_settings
numeric_params <- c("alpha", "alpha_d", "ovp", "dovp", "C_target", "C_target_pen", "D_target", "D_target_pen", "pred_days", "sim_ens", "ndays", "N")
sim_settings[numeric_params] <- lapply(sim_settings[numeric_params], as.numeric)

#Set this up for parallel run

# Create and export cluster
Ncores <-  2  # Adjust cores as appropriate
cl <- makeCluster(Ncores)

clusterExport(cl, ls())
clusterExport(cl, c("reward_fun_wd", "Epi_pred_wd", "Epi_MPC_run_wd", "episim_data_ens", "episettings", "Action_space", "Epi_pars", "Noise_pars", "R_estim"))

parallel::clusterEvalQ(cl, {
  library(pbapply)
  library(VGAM)
  library(EpiControl)
})

episettings$parallel <- TRUE
episettings$cl <- cl

# Run the epicontrol function
results <- epicontrol(episim_data_ens, episettings)

# Stop the cluster after simulation
parallel::stopCluster(cl)

```

### Results

Plot the results

```{r plotting}

# Combine Simulation Results
combined_data <- do.call(rbind, results)

combined_data <- combined_data %>%
  group_by(sim_id, policy) %>%
  arrange(days) %>%
  mutate(group = cumsum(c(1, diff(days) != 1))) %>%
  ungroup()

combined_data <- combined_data %>%
  arrange(sim_id, days, policy)

policy_labels <- c("1" = "No intervention", "2" = "Social distancing", "3" = "Lockdown")


ggplot(combined_data %>% filter(sim_id == 1)) +
  geom_line(data = subset(combined_data, sim_id != 1), aes(x = days, y = C, color = as.factor(sim_id)), alpha = 0.1) +
  geom_line(aes(x = days, y = C, color = factor(policy, labels = policy_labels), group = 1), alpha = 1.0) +
  geom_line(aes(x = days, y = C, color = factor(policy, labels = policy_labels), group = 1), alpha = 1.0, linewidth=0.25) +
  geom_hline(yintercept = C_target, linetype = "dashed", color = "blue", linewidth=0.25) +
  labs(x = "Days", y = "Reported cases", color = "Policy") +
  scale_color_manual(values = c("No intervention" = "chartreuse3", "Social distancing" = "violet", "Lockdown" = "red")) +
  guides(color = guide_legend(title = "Policy"))

```
