---
title: "Simulating Epidemic Dynamics with Policy Interventions"
author: "Sandor Beregi"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo=FALSE, results='asis'}
cat("## Overview\n\n",
    "This vignette demonstrates model predictive optimal control of ICU cases for COVID-19 \n\n")
```

### Load Libraries and Setup

Load the required libraries and initialize the simulation environment.

```{r load-libraries}
library(VGAM)
library(parallel)
library(pbapply)
library(zoo)  # For rolling sum operations
library(dplyr)
library(ggplot2)
```

Initialize the cluster for parallel computing:

```{r parallel-setup}
cores <- detectCores() - 1
cl <- makeCluster(cores)
```

### Epidemiological and Noise Parameters

Define parameters for pathogens and noise:

```{r define-parameters}
Epi_pars <- data.frame(
  Pathogen = c("COVID-19", "Ebola"),
  R0 = c(2.2, 2.5),
  gen_time = c(6.5, 15.0),
  gen_time_var = c(2.1, 2.1),
  CFR = c(0.0132, 0.5),
  mortality_mean = c(10.0, 10.0),
  mortality_var = c(1.1, 1.1)
)

Noise_pars <- data.frame(
  repd_mean = 10.5,  # Reporting delay mean
  del_disp = 5.0,    # Reporting delay variance
  ur_mean = 0.3,     # Under-reporting mean
  ur_beta_a = 50.0   # Beta distribution alpha for under-reporting
)
```

### Action Space for Policy Interventions

Define non-pharmaceutical interventions:

```{r action-space}
Action_space <- data.frame(
  NPI = c("No restrictions", "Lockdown"),
  R_coeff = c(1.0, 0.3),
  R_beta_a = c(0.0, 5.0),
  cost_of_NPI = c(0.0, 0.15)
)
```

### Simulation Setup

Set up key parameters and initialize simulation data:

```{r simulation-setup}
ndays <- 83 * 7
N <- 1e7
I0 <- 10

column_names <- c("days", "sim_id", "I", "Lambda", "C", "Lambda_C", "S", "Deaths", "Re", "Rew", "Rest", "R0est", "policy", "R_coeff")
episim_data <- data.frame(matrix(0, nrow = ndays, ncol = length(column_names)))
colnames(episim_data) <- column_names

episim_data$policy <- rep(1, ndays)
episim_data$days <- 1:ndays
episim_data[1, ] <- c(1, 1, I0, I0, Noise_pars$ur_mean * I0, Noise_pars$ur_mean * I0, N - I0, 0, Epi_pars[1, "R0"], Epi_pars[1, "R0"], 1, 1, 1)
```

### Running the Simulation

Run simulations in parallel using the pblapply function:

```{r run-simulation}
episim_data_ens <- replicate(100, episim_data, simplify = FALSE)

clusterExport(cl, c("episim_data_ens", "Epi_pars", "Noise_pars", "Action_space", "logistic_function", "Epi_MPC_run_wd"))

results <- pblapply(1:100, function(idx) {
  Epi_MPC_run_wd(episim_data_ens[[idx]], Epi_pars, Noise_pars, Action_space, pred_days = 28, n_ens = 100, start_day = 1, ndays = ndays, R_est_wind = 5, pathogen = 1, susceptibles = 0, delay = 1, ur = 1, r_dir = 1, N = N)
}, cl = cl)

stopCluster(cl)

# Combine Simulation Results
combined_data <- do.call(rbind, results)
```
